\vskip-0.6in Temporary Trade Barriers: When Will They End?

### 2017-09-18 draft.tex
add game-tree.tex, modify title/abstract, remove pieces necessary for first compilation

\begin{document}
\title{\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}}
\thanksmarkseries{alph}
\author{Kristy Buzard\thanks{Syracuse University, Economics Department, 110 Eggers Hall, Syracuse, NY 13244, USA. Ph: 315-443-4079. Fax: 315-443-3717. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard. I gratefully acknowledge funding from the W. Allen Wallis Institute of Political Economy.}} 
\date{\vskip-.1in \today}
\maketitle

@@

\begin{quote}
{This paper presents a model of international trade agreements in which the executive branches of each government negotiate agreements while the legislative branches, subject to political pressure from firms, can disrupt them. Lobbying is in the style of Grossman and Helpman (1994) with a new feature: all actors face uncertainty arising from the complexity of the legislative process. I demonstrate that the higher the executives set tariffs in a trade agreement, the less effort lobbies put forth to prevent its ratification. Thus trade agreements act as a domestic political commitment device: executives set relatively high tariffs to discourage lobbying and increase the chance that the agreement will be ratified. The model sheds light on the empirical puzzle surrounding governments' welfare weights in the Grossman and Helpman (1994) model and provides a new explanation for failures to ratify trade agreements.

\textit{JEL classification:} D72, D80, F13, F53, F55 \\
\textit{Keywords:} trade agreements, protection, lobbying, political economy, uncertainty}
\end{quote}

@@

%http://www.whitehouse.gov/sites/default/files/fact_sheet_increasing_us_auto_exports_us_korea_free_trade_agreement.pdf

@@

\begin{figure}
\begin{center}
\includegraphics[height=2.5in, width=3in]{brprob2.jpg}
\end{center}
\caption{Probability Trade Agreement will be Broken\label{fig:br}}
\end{figure}

@@

\begin{comment}
\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{brprob2.jpg}
  \caption{Probability Trade Agreement will be Broken}
  \label{fig:br}
  %\label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{contributions.jpg}
  \caption{Lobbying Effort}
  \label{fig:cont}
  %\label{fig:sub2}
\end{subfigure}
%\caption{A figure with two subfigures}
%\label{fig:test}
\end{figure}
\end{comment}

Figure~\ref{fig:br} depicts the probability that the legislature will vote to break the trade agreement as a function of the tariff levels set in the trade agreement (with the restriction that $\tau^a = \tau^{*a}$) and the lobby's effort. It is strongly increasing in lobbying effort and decreasing in the level of tariffs set in the trade agreements.

Given the impact of lobbying on the legislature's break decision, the lobby's optimal contribution level is strongly decreasing in the trade agreement tariffs.%, as shown in Figure~\ref{fig:cont}.

@@

\begin{figure}
\begin{center}
\includegraphics[height=2.5in, width=3in]{lobby_br.jpg}
\end{center}
\caption{Lobby's Optimal Expenditure Function (Low and Intermediate Uncertainty)\label{fig:lobby_br}}
\end{figure}

Figure~\ref{fig:lobby_br} shows the lobby's reaction function at this very low level of uncertainty as well as the intermediate level when $\ve$ is distributed uniformly on $[-0.14,0.14]$ where the executives set the highest tariff level of $0.107$. For the former, the range of tariff levels to which the lobby responds with a contribution that leaves open the possibility of a break is very small: only those between 0.103 and 0.104. The lobby disengages completely at $\bta=0.105$. When uncertainty is increased to the interval $[-0.14,0.14]$, the lobby begins to leave open the possibility of a break at $\bta = 0.048$. It disengages at the optimal tariff level of $0.107$. At intermediate uncertainly levels, both portions of the best response curve are steeper and the lobby disengages at lower tariff levels---as low as $0.099$ when $\ve \sim U[-0.06,0.06]$.  The optimal tariff does not decrease monotonically.

This illustrates Part (a) of Result~\ref{res:execsoln}: here the executives maximize their welfare by raising tariffs to the point where the import-competing industry ceases to lobby and the agreement remains in force for sure. Although lobbying expenditures are zero, the potential for lobbying behavior is essential in shaping the trade agreement.

%\begin{figure}
%\begin{center}
%\includegraphics[height=3.8in, width=7in]{exec_welfare.jpg}
%\end{center}
%\caption{Differences in Executive Welfare when Uncertainty Increases\label{fig:exec_wel}}
%\end{figure}

When uncertainty rises above this threshold of $\ve \sim U[-0.14,0.14]$, the executives' choices are made according to Part (b) of Result~\ref{res:execsoln}: they trade off reductions in welfare under the agreement with reductions in the probability that the agreement will be abrogated by the legislature. %In terms of the lobby's best response function, the executives now choose the optimal point on the downward-sloping portion of the curve instead of the point where it reaches zero. In terms of the welfare function for the executives, shown in Figure~\ref{fig:exec_wel}, they are now choosing the point that maximizes the concave portion of the curve instead of the ``spike'' that is created when the lobbies disengage and the chance that the agreement will be abrogated is removed altogether.\footnote{Note that the initial, flat portion of the curve is where the tariff is low enough that the lobby buys a trade war with certainty; we then enter the concave portion where the main results of Section~\ref{sec:main} are applicable; finally, we see the upward spike where the probability of a trade war is reduced to zero and the executives' welfare declines after that because there is no further reduction in break probability as the tariff increases.}

We see from this example that variations in uncertainty lead to very different outcomes in terms of lobbying, trade agreement tariffs and the probability of disruption. The provided tariff levels are strongly influenced by lobbying incentives, but in more nuanced ways than those predicted by unitary models.

>>>>

\begin{document}
\title{\vskip-0.6in Temporary Trade Barriers: When Will They End?}%\protect\footnote{Formerly circulated under the title }}
\thanksmarkseries{alph}
\author{Kristy Buzard\thanks{Syracuse University, Economics Department, 110 Eggers Hall, Syracuse, NY 13244, USA. Ph: 315-443-4079. Fax: 315-443-3717. Email: kbuzard@syr.edu. http://faculty.maxwell.syr.edu/kbuzard.}} 
\date{\vskip-.1in \today}
\maketitle

@@

\begin{quote}
{Abstract

\textit{JEL classification:} D72, F13, F55 \\
\textit{Keywords:} temporary trade barriers, WTO, disputes, lobbying, political economy}
\end{quote}

@@

@@

@@

@@
